---
title: "SSN703 Rating Curve Development -- v5"
author: "[Emily Haughton]"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

library(tidyverse)
library(plotly)
library(segmented)
library(stats)
library(KScorrect)

select <- dplyr::select

source("02_processing/scripts/selectspan.R")
source("02_processing/scripts/interp1.R")
source("02_processing/scripts/interp2.R")
source("02_processing/scripts/HQ_unc.R")
source("02_processing/scripts/CI_model_input.R")
source("02_processing/scripts/CI.R")

out_dir <- "04_outputs"
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

n_boot <- 500
```

```{r load_data, include=FALSE}
# NOTE: Stage_avg = Stage_avg_corrected throughout this script
# renamed for compatibility with colleague's source functions

gaugings_all <- read_csv("03_docs/metadata/ssn703_gaugings_prepped.csv") |>
  filter(!is.na(Q_meas)) |>
  mutate(
    Stage_avg = Stage_avg_corrected,
    Q_abs_unc = Q_rel_unc / 100 * Q_meas,
    datetime  = as.Date(datetime)
  )

# Previous lookup table -- for context/diagnostic only
# Discharge record pre-2019-02-09 is already in the database and not touched
ratingcurve_v4 <- read_csv("03_docs/metadata/ssn703_RC1_lookup_v_previous.csv")
rc_v4_r3 <- ratingcurve_v4 |> filter(Rating == 3) |> dplyr::select(Stage_avg, Q_model)

# RC2 -- loc_2, sensor C
# Start date for applied curve: 2019-02-09 (last date covered by previous pipeline)
# All gaugings with ok stage; exclude WY 2020-2021 (sparse/biased)
HQ_RC2 <- gaugings_all |>
  filter(rating_curve_period == "RC2",
         stage_status == "ok",
         WY != "2020-2021",
         Old_New == "New",
         Final_rating_curve == "Y")

HQ_RC2_old <- HQ_RC2 |> filter(Old_New == "Old")
HQ_RC2_new <- HQ_RC2 |> filter(Old_New == "New")

# RC3 -- loc_3, sensor D; all New
# Event 63 rising limb at ~121cm excluded (physically implausible vs neighbours)
HQ_RC3 <- gaugings_all |>
  filter(rating_curve_period == "RC3",
         stage_status == "ok",
         Old_New == "New",
         Final_rating_curve == "Y") |>
  filter(!(Event_no == 63 & Stage_avg > 120 & Stage_avg < 122))
```

---

## Background

The discharge record for SSN703 up to **2019-02-09** is already computed and stored in the database using the previous pipeline (v4, Ratings 1-3). That record is not touched here.

This document develops two new rating curves covering the period from 2019-02-09 onward:

| Rating | Sensor | Location | Applied from | Notes |
|:-------|:-------|:---------|:-------------|:------|
| RC2 | ssn703_c | loc_2 | 2019-02-09 | New location; previous pipeline never incorporated sensor C |
| RC3 | ssn703_d | loc_3 | 2023-09-15 | New location; sensor D installed 2021-09-02 |

**Key decisions:**

- Stage_avg throughout = Stage_avg_corrected from the prepped gauging table
- WY 2020-2021 excluded from RC2 due to sparse and biased stage coverage
- Stage_suspect window (2023-06-25 to 2023-09-15) excluded -- ssn703_c stage bad before db updated to ssn703_d
- Old/New in plots: Old = gaugings collected before 2019-02-09; New = post 2019-02-09
- RC3 Event 63 rising limb gauging at ~121cm excluded (physically implausible vs neighbours)

---

## Diagnostic -- RC2 gaugings vs previous Rating 3 curve

_Early RC2 gaugings (loc_2, sensor C) plotted against the last active previous curve (v4 Rating 3, loc_1 sensor B). This confirms that loc_2 produces a systematically different stage-discharge relationship and justifies starting a new curve._

```{r diagnostic}
HQ_RC2_diag <- gaugings_all |>
  filter(rating_curve_period == "RC2",
         stage_status == "ok",
         Old_New == "New")

ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r3,
              aes(x = Stage_avg, y = Q_model),
              colour = "black", linetype = "dashed") +
    geom_pointrange(data = HQ_RC2_diag,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        colour = Final_rating_curve,
                        label = paste0("Event: ", Event_no,
                                       " | Date: ", datetime,
                                       " | Q: ", round(Q_meas, 3))),
                    size = 0.5) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 gaugings vs previous Rating 3 curve",
         subtitle = "Dashed = v4 Rating 3 (last active previous curve)",
         colour = "Final rating curve",
         caption = "Y = used for fitting; N = excluded")
)
```

---

## RC2 -- loc_2 (2019-02-09 onward) {.tabset}

_`r nrow(HQ_RC2)` gaugings used for RC2 (`r nrow(HQ_RC2_old)` Old / `r nrow(HQ_RC2_new)` New). No previous curve for this location. WY 2020-2021 excluded._

### Data assessment {.tabset}

#### Overview

```{r rc2_overview}
ggplotly(
  ggplot() +
    geom_point(data = HQ_RC2,
               aes(x = Stage_avg, y = Q_meas,
                   fill = Old_New,
                   label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 -- all gaugings (Old vs New)",
         fill = "Old/New",
         caption = "Old = available to previous pipeline but unused; New = post 2019-02-09")
)
```

#### By water year

```{r rc2_wy}
ggplotly(
  ggplot() +
    geom_point(data = HQ_RC2,
               aes(x = Stage_avg, y = Q_meas,
                   fill = factor(WY),
                   label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 -- gaugings by water year", fill = "Water year")
)
```

#### With uncertainty

```{r rc2_unc}
ggplotly(
  ggplot() +
    geom_pointrange(data = HQ_RC2,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        colour = Old_New,
                        label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                       " | Q: ", round(Q_meas, 3))),
                    size = 0.5) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 -- gaugings with uncertainty", colour = "Old/New")
)
```

#### By event

```{r rc2_event}
ggplotly(
  ggplot() +
    geom_point(data = HQ_RC2,
               aes(x = Stage_avg, y = Q_meas,
                   fill = factor(Event_no),
                   label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 -- gaugings by event", fill = "Event no.")
)
```

#### Hysteresis

```{r rc2_hyst}
ggplotly(
  ggplot() +
    geom_point(data = HQ_RC2,
               aes(x = Stage_avg, y = Q_meas,
                   fill = Stage_delta,
                   label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 -- hysteresis assessment", fill = "Stage delta (R/F/C)")
)
```

### LOESS span selection

```{r rc2_span}
Rating_RC2 <- selectspan(HQ_RC2, 0.3, 0.4, 0.5)

ggplotly(
  ggplot() +
    geom_pointrange(data = HQ_RC2,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        label = Event_no),
                    size = 0.5) +
    geom_line(data = Rating_RC2,
              aes(x = Stage_avg, y = Q_model, colour = smoothed)) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 -- LOESS span selection")
)
```

```{r rc2_span_choose, include=FALSE}
## ACTION REQUIRED: enter chosen span label and value
RC2_span_label <- "span_b"
RC2_span_value <- 0.4

Rating_RC2 <- Rating_RC2 |>
  filter(smoothed == RC2_span_label) |>
  dplyr::select(Stage_avg, Q_model)
```

### Extrapolation and final curve

```{r rc2_extrap, include=FALSE}
h0_RC2 <- min(HQ_RC2$Stage_avg[HQ_RC2$Q_meas > 0], na.rm = TRUE) * 0.9

## ACTION REQUIRED: set inflection point manually
## Power law fitted only on gaugings at or above this stage
inflection_h_RC2 <- 150

# Derive starting values from log-log linear model on upper data
RC2_extrap_data <- HQ_RC2 |> filter(Stage_avg >= inflection_h_RC2, Q_meas > 0)
RC2_loglog      <- RC2_extrap_data |>
  mutate(log_h = log(Stage_avg - h0_RC2), log_Q = log(Q_meas))
lm_start_RC2 <- lm(log_Q ~ log_h, data = RC2_loglog)
a_start_RC2  <- exp(coef(lm_start_RC2)[1])
b_start_RC2  <- coef(lm_start_RC2)[2]

fit_RC2 <- tryCatch(
  nls(Q_meas ~ a * (Stage_avg - h0)^b,
      data    = RC2_extrap_data,
      start   = list(a = a_start_RC2, b = b_start_RC2, h0 = h0_RC2),
      control = nls.control(maxiter = 500)),
  error = function(e) {
    message("RC2 nls failed with free h0 -- falling back to fixed h0")
    nls(Q_meas ~ a * (Stage_avg - h0_RC2)^b,
        data    = RC2_extrap_data,
        start   = list(a = a_start_RC2, b = b_start_RC2),
        control = nls.control(maxiter = 500))
  }
)

# Anchor power law to last LOESS value so the join is smooth
RC2_max_loess    <- max(Rating_RC2$Stage_avg)
RC2_Q_at_anchor  <- Rating_RC2 |>
  filter(Stage_avg == RC2_max_loess) |>
  pull(Q_model)

# Rescale power law predictions to pass through the LOESS anchor point
RC2_Q_pl_at_anchor <- predict(fit_RC2,
                               newdata = data.frame(Stage_avg = RC2_max_loess))
RC2_pl_scale <- RC2_Q_at_anchor / RC2_Q_pl_at_anchor

RC2_extrap_stage  <- seq(RC2_max_loess + 1, 218, by = 1)
RC2_extrap_Q      <- predict(fit_RC2,
                              newdata = data.frame(Stage_avg = RC2_extrap_stage)) * RC2_pl_scale
RC2_extrap_values <- tibble(Stage_avg = RC2_extrap_stage, Q_model = RC2_extrap_Q)
Rating_RC2_full   <- bind_rows(Rating_RC2, RC2_extrap_values) |> arrange(Stage_avg)

RC2_params <- coef(fit_RC2)
```

_RC2 h₀ = `r round(h0_RC2, 1)` cm | Inflection = `r round(inflection_h_RC2, 1)` cm | Power law anchored at LOESS endpoint (`r round(RC2_max_loess, 1)` cm, `r round(RC2_Q_at_anchor, 2)` m³/s) | a = `r round(RC2_params["a"], 4)`, b = `r round(RC2_params["b"], 4)`_

```{r rc2_extrap_plot}
ggplotly(
  ggplot() +
    geom_vline(xintercept = inflection_h_RC2, linetype = "dotted", colour = "grey50") +
    geom_line(data = Rating_RC2_full,
              aes(x = Stage_avg, y = Q_model,
                  colour = Stage_avg > RC2_max_loess)) +
    geom_point(data = HQ_RC2,
               aes(x = Stage_avg, y = Q_meas,
                   label = paste0("Event: ", Event_no, " | Date: ", datetime)),
               colour = ifelse(HQ_RC2$Stage_avg >= inflection_h_RC2, "#E07B54", "black"),
               size = 2) +
    scale_colour_manual(values = c("FALSE" = "black", "TRUE" = "red"),
                        labels = c("LOESS", "Power law extrapolation")) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC2 -- LOESS fit with anchored power law extrapolation",
         subtitle = paste0("Dotted line = inflection (", round(inflection_h_RC2, 1),
                           " cm). Orange = used for power law fit. Power law rescaled to LOESS at ",
                           round(RC2_max_loess, 1), " cm."),
         colour = "Fit type")
)
```

```{r rc2_interp, include=FALSE}
Rating_RC2_interp <- interp1(Rating_RC2_full)
names(Rating_RC2_interp) <- c("Stage_avg", "Q_model")
```

```{r rc2_ci, include=FALSE}
HQ_RC2_unc   <- HQ_unc(HQ_RC2, Rating_RC2, Rating_RC2_interp)
CI_input_RC2 <- CI_model_input(HQ_RC2_unc, n_boot, HQ_RC2)
CI_RC2       <- CI(CI_input_RC2, n_boot, RC2_span_value, 1)
RatingCI_RC2 <- bind_cols(Rating_RC2, CI_RC2)

RC2_extrap_CI <- RC2_extrap_values |>
  mutate(Max_CI = Q_model * 1.05, Min_CI = Q_model * 0.95)

RatingCI_RC2_full   <- bind_rows(RatingCI_RC2, RC2_extrap_CI) |> arrange(Stage_avg)
RatingCI_RC2_interp <- interp2(RatingCI_RC2_full)
```

```{r rc2_final_plot}
Q_meas_RC2 <- cbind(HQ_RC2$Stage_avg, HQ_RC2_unc)
colnames(Q_meas_RC2)[1] <- "Stage_avg"

ggplotly(
  ggplot() +
    geom_ribbon(data = RatingCI_RC2_interp,
                aes(x = Stage_avg, ymin = Min_CI, ymax = Max_CI),
                fill = "grey80") +
    geom_pointrange(data = Q_meas_RC2,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_min, ymax = Q_max),
                    size = 0.5) +
    geom_line(data = RatingCI_RC2_interp,
              aes(x = Stage_avg, y = Q_model),
              colour = "#333333") +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN703 RC2 -- Final rating curve v1")
)
```

```{r rc2_save, include=FALSE}
RC2_lookup <- RatingCI_RC2_interp |>
  mutate(
    Start  = format(as.Date("2019-02-09"), "%m/%d/%Y %H:%M"),
    Rating = 4L
  ) |>
  dplyr::select(Start, Stage_avg, Q_model, Max_CI, Min_CI, Rating)

write_csv(RC2_lookup, file.path(out_dir, "ssn703_RC2_rating_curve_v1.csv"))
write_csv(HQ_RC2,    file.path(out_dir, "ssn703_RC2_stagedischarge_v1.csv"))
```

---

## RC3 -- loc_3 (2023-09-15 onward) {.tabset}

_`r nrow(HQ_RC3)` gaugings used for RC3. All New -- sensor D at loc_3 entirely post-dates the previous pipeline. No previous curve for this location. **Caution: sparse data.**_

### Data assessment {.tabset}

#### Overview

```{r rc3_overview}
ggplotly(
  ggplot() +
    geom_point(data = HQ_RC3,
               aes(x = Stage_avg, y = Q_meas,
                   fill = factor(WY),
                   label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC3 -- all gaugings by water year",
         fill = "Water year",
         caption = "Caution: sparse data")
)
```

#### With uncertainty

```{r rc3_unc}
ggplotly(
  ggplot() +
    geom_pointrange(data = HQ_RC3,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        colour = factor(WY),
                        label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                       " | Q: ", round(Q_meas, 3))),
                    size = 0.5) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC3 -- gaugings with uncertainty by water year",
         colour = "Water year")
)
```

#### Hysteresis

```{r rc3_hyst}
ggplotly(
  ggplot() +
    geom_point(data = HQ_RC3,
               aes(x = Stage_avg, y = Q_meas,
                   fill = Stage_delta,
                   label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC3 -- hysteresis assessment", fill = "Stage delta (R/F/C)")
)
```

### LOESS span selection

```{r rc3_span}
Rating_RC3 <- selectspan(HQ_RC3, 0.5, 0.6, 0.7)

ggplotly(
  ggplot() +
    geom_pointrange(data = HQ_RC3,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        label = Event_no),
                    size = 0.5) +
    geom_line(data = Rating_RC3,
              aes(x = Stage_avg, y = Q_model, colour = smoothed)) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC3 -- LOESS span selection (retained for documentation only)",
         caption = "Final curve uses full power law -- see below")
)
```

```{r rc3_span_choose, include=FALSE}
## Retained for documentation -- LOESS not used for final RC3 curve
RC3_span_label <- "span_b"
RC3_span_value <- 0.60

Rating_RC3 <- Rating_RC3 |>
  filter(smoothed == RC3_span_label) |>
  dplyr::select(Stage_avg, Q_model)
```

```{r rc3_loess_full, include=FALSE}
# Build full LOESS curve for comparison plot
RC3_loess_stage <- seq(min(HQ_RC3$Stage_avg), max(HQ_RC3$Stage_avg), by = 0.5)
RC3_loess_Q     <- predict(loess(Q_meas ~ Stage_avg, data = HQ_RC3,
                                  span = RC3_span_value),
                            newdata = data.frame(Stage_avg = RC3_loess_stage))
Rating_RC3_full <- tibble(Stage_avg = RC3_loess_stage, Q_model = RC3_loess_Q)
```

### Power law fit comparison

_Given sparse data, a fully parametric power law fit is compared against the LOESS._

```{r rc3_powerlaw}
h0_RC3_pl <- min(HQ_RC3$Stage_avg[HQ_RC3$Q_meas > 0], na.rm = TRUE) * 0.9

RC3_pl_loglog <- HQ_RC3 |>
  filter(Stage_avg > h0_RC3_pl, Q_meas > 0) |>
  mutate(log_h = log(Stage_avg - h0_RC3_pl), log_Q = log(Q_meas))

lm_pl_RC3  <- lm(log_Q ~ log_h, data = RC3_pl_loglog)
a_pl_start <- exp(coef(lm_pl_RC3)[1])
b_pl_start <- coef(lm_pl_RC3)[2]

fit_RC3_pl <- tryCatch(
  nls(Q_meas ~ a * (Stage_avg - h0)^b,
      data    = HQ_RC3 |> filter(Stage_avg > h0_RC3_pl, Q_meas > 0),
      start   = list(a = a_pl_start, b = b_pl_start, h0 = h0_RC3_pl),
      control = nls.control(maxiter = 500)),
  error = function(e) {
    message("RC3 power law failed with free h0 -- falling back to fixed h0")
    nls(Q_meas ~ a * (Stage_avg - h0_RC3_pl)^b,
        data    = HQ_RC3 |> filter(Stage_avg > h0_RC3_pl, Q_meas > 0),
        start   = list(a = a_pl_start, b = b_pl_start),
        control = nls.control(maxiter = 500))
  }
)

RC3_pl_stage  <- seq(min(HQ_RC3$Stage_avg), 218, by = 0.1)
RC3_pl_Q      <- predict(fit_RC3_pl, newdata = data.frame(Stage_avg = RC3_pl_stage))
RC3_pl_curve  <- tibble(Stage_avg = RC3_pl_stage, Q_model = RC3_pl_Q)
RC3_pl_params <- coef(fit_RC3_pl)

ggplotly(
  ggplot() +
    geom_pointrange(data = HQ_RC3,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        label = paste0("Event: ", Event_no, " | Date: ", datetime)),
                    size = 0.5) +
    geom_line(data = Rating_RC3_full,
              aes(x = Stage_avg, y = Q_model, colour = "LOESS"),
              linewidth = 1) +
    geom_line(data = RC3_pl_curve,
              aes(x = Stage_avg, y = Q_model, colour = "Full power law"),
              linewidth = 1, linetype = "dashed") +
    scale_colour_manual(values = c("LOESS" = "black", "Full power law" = "#E07B54")) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "RC3 -- LOESS vs full power law fit",
         subtitle = paste0("Power law: a = ", round(RC3_pl_params["a"], 4),
                           ", b = ", round(RC3_pl_params["b"], 4),
                           ", h0 = ", round(RC3_pl_params["h0"], 1), " cm"),
         colour = "Fit type")
)
```

### Power law fit -- final curve

_Full power law Q = a(h - h0)^b fitted on all RC3 gaugings. CI = ±10% throughout._

```{r rc3_final, include=FALSE}
h0_RC3_pl <- min(HQ_RC3$Stage_avg[HQ_RC3$Q_meas > 0], na.rm = TRUE) * 0.9

RC3_pl_loglog <- HQ_RC3 |>
  filter(Stage_avg > h0_RC3_pl, Q_meas > 0) |>
  mutate(log_h = log(Stage_avg - h0_RC3_pl), log_Q = log(Q_meas))

lm_pl_RC3  <- lm(log_Q ~ log_h, data = RC3_pl_loglog)
a_pl_start <- exp(coef(lm_pl_RC3)[1])
b_pl_start <- coef(lm_pl_RC3)[2]

fit_RC3_pl <- tryCatch(
  nls(Q_meas ~ a * (Stage_avg - h0)^b,
      data    = HQ_RC3 |> filter(Stage_avg > h0_RC3_pl, Q_meas > 0),
      start   = list(a = a_pl_start, b = b_pl_start, h0 = h0_RC3_pl),
      control = nls.control(maxiter = 500)),
  error = function(e) {
    message("RC3 power law failed with free h0 -- falling back to fixed h0")
    nls(Q_meas ~ a * (Stage_avg - h0_RC3_pl)^b,
        data    = HQ_RC3 |> filter(Stage_avg > h0_RC3_pl, Q_meas > 0),
        start   = list(a = a_pl_start, b = b_pl_start),
        control = nls.control(maxiter = 500))
  }
)

RC3_pl_params <- coef(fit_RC3_pl)
a_RC3_pl <- RC3_pl_params["a"]
b_RC3_pl <- RC3_pl_params["b"]
h0_RC3   <- RC3_pl_params["h0"]

# Extend curve down to h0 + 1cm -- power law naturally approaches zero at h0
# This is extrapolation below the minimum gauged stage (56.88cm) and carries
# higher uncertainty than the fitted range. Widened CI (±20%) applied below
# minimum gauged stage to reflect absence of gauging constraint at low flows.
# The PT records stage below 56.88cm during dry summer periods -- without this
# extrapolation those timesteps would return NA discharge. No low flow gaugings
# exist at loc_3 to anchor the curve at the low end; this should be a priority
# for future field seasons.

RC3_min_gauged <- min(HQ_RC3$Stage_avg)
RC3_pl_stage   <- seq(round(h0_RC3 + 1, 1), 218, by = 0.1)
RC3_pl_Q       <- predict(fit_RC3_pl, newdata = data.frame(Stage_avg = RC3_pl_stage))
RC3_pl_curve   <- tibble(Stage_avg = RC3_pl_stage, Q_model = RC3_pl_Q)

RC3_pl_CI <- RC3_pl_curve |>
  mutate(
    Max_CI = ifelse(Stage_avg < RC3_min_gauged,
                    Q_model * 1.20,   # wider CI below minimum gauged stage
                    Q_model * 1.10),
    Min_CI = ifelse(Stage_avg < RC3_min_gauged,
                    Q_model * 0.80,
                    Q_model * 0.90)
  )
```

_RC3 power law: a = `r round(a_RC3_pl, 4)`, b = `r round(b_RC3_pl, 4)`, h₀ = `r round(h0_RC3, 1)` cm. Curve extended from h₀+1cm to 218cm. CI = ±10% within gauged range (`r round(RC3_min_gauged, 1)`--218cm); ±20% below minimum gauged stage (extrapolated low flow range -- **no gaugings exist below `r round(RC3_min_gauged, 1)`cm at loc_3**)._

```{r rc3_final_plot}
ggplotly(
  ggplot() +
    geom_ribbon(data = RC3_pl_CI,
                aes(x = Stage_avg, ymin = Min_CI, ymax = Max_CI),
                fill = "grey80") +
    geom_vline(xintercept = RC3_min_gauged, linetype = "dashed",
               colour = "orange") +
    geom_pointrange(data = HQ_RC3,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        label = paste0("Event: ", Event_no, " | Date: ", datetime,
                                       " | Q: ", round(Q_meas, 3))),
                    size = 0.5) +
    geom_line(data = RC3_pl_curve,
              aes(x = Stage_avg, y = Q_model,
                  colour = Stage_avg < RC3_min_gauged)) +
    scale_colour_manual(values = c("FALSE" = "#333333", "TRUE" = "#E07B54"),
                        labels = c("FALSE" = "Fitted range",
                                   "TRUE"  = "Extrapolated (no gaugings)")) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN703 RC3 -- Final rating curve v1 (power law)",
         subtitle = paste0("Orange dashed = minimum gauged stage (",
                           round(RC3_min_gauged, 1),
                           "cm). Orange line = extrapolated low flow range (±20% CI). ",
                           "Caution: provisional curve, sparse gaugings."),
         colour = "Curve range",
         caption = "Caution: sparse data -- provisional curve, to be updated as gaugings accumulate")
)
```

```{r rc3_save, include=FALSE}
RC3_lookup <- RC3_pl_CI |>
  mutate(
    Start  = format(as.Date("2023-09-15"), "%m/%d/%Y %H:%M"),
    Rating = 5L
  ) |>
  dplyr::select(Start, Stage_avg, Q_model, Max_CI, Min_CI, Rating)

write_csv(RC3_lookup, file.path(out_dir, "ssn703_RC3_rating_curve_v1.csv"))
write_csv(HQ_RC3,    file.path(out_dir, "ssn703_RC3_stagedischarge_v1.csv"))

# Combined lookup -- append new ratings to existing v4 lookup
ssn703_lookup_combined <- bind_rows(
  ratingcurve_v4 |> dplyr::select(Start, Stage_avg, Q_model, Max_CI, Min_CI, Rating),
  RC2_lookup,
  RC3_lookup
)
write_csv(ssn703_lookup_combined,
          file.path(out_dir, "ssn703_rating_curve_lookup_combined.csv"))
```

---

## Summary

| Rating | Period | n gaugings | Stage range (cm) | h₀ (cm) | Span | Version |
|:-------|:-------|:-----------|:-----------------|:---------|:-----|:--------|
| RC2 (Rating 4) | 2019-02-09 to 2023-06-25 | `r nrow(HQ_RC2)` | `r round(min(HQ_RC2$Stage_avg),1)` -- `r round(max(HQ_RC2$Stage_avg),1)` | `r round(h0_RC2,1)` | `r RC2_span_value` | v1 |
| RC3 (Rating 5) | 2023-09-15 onward | `r nrow(HQ_RC3)` | `r round(min(HQ_RC3$Stage_avg),1)` -- `r round(max(HQ_RC3$Stage_avg),1)` | `r round(h0_RC3,1)` | Power law | v1 |

**Previous pipeline (Ratings 1-3) covers up to 2019-02-09 and is not modified.**

**Output files written to `04_outputs/`:**

- `ssn703_RC2_rating_curve_v1.csv` -- Rating 4, loc_2
- `ssn703_RC3_rating_curve_v1.csv` -- Rating 5, loc_3
- `ssn703_RC2_stagedischarge_v1.csv` -- RC2 gauging table
- `ssn703_RC3_stagedischarge_v1.csv` -- RC3 gauging table
- `ssn703_rating_curve_lookup_combined.csv` -- v4 Ratings 1-3 plus new Ratings 4-5
