---
title: "SSN844 Rating Curve Development"
author: "[your name]"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

library(tidyverse)
library(plotly)
library(segmented)

source("02_processing/scripts/selectspan.R")
source("02_processing/scripts/interp1.R")
source("02_processing/scripts/interp2.R")
source("02_processing/scripts/HQ_unc.R")
source("02_processing/scripts/CI_model_input.R")
source("02_processing/scripts/CI.R")

select <- dplyr::select

out_dir <- "04_outputs"
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

n_boot <- 500
```

```{r load_data, include=FALSE}
gaugings_all <- read_csv("03_docs/metadata/ssn844_gaugings_prepped.csv") |>
  filter(!is.na(Q_meas)) |>
  mutate(
    Stage_avg = Stage_avg_corrected,
    Q_abs_unc = Q_rel_unc / 100 * Q_meas,
    datetime  = as.Date(datetime)
  )

# Previous lookup table
ratingcurve_v4 <- read_csv("03_docs/metadata/ssn844_RC1_lookup_v_previous.csv")

rc_v4_r1 <- ratingcurve_v4 |> filter(Rating == 1) |> dplyr::select(Stage_avg, Q_model)
rc_v4_r2 <- ratingcurve_v4 |> filter(Rating == 2) |> dplyr::select(Stage_avg, Q_model)

# All gaugings with ok stage
gaugings_ok <- gaugings_all |> filter(stage_status == "ok")

# Split by Old_New for diagnostic
gaugings_old <- gaugings_ok |> filter(Old_New == "Old")
gaugings_new <- gaugings_ok |> filter(Old_New == "New"& Final_rating_curve=="Y")
```

---

## Site background

SSN844 has a single pressure transducer (`ssn844`) at `loc_1` from 2014-09-10 to present. No sensor location changes — only one physical location throughout the record.

The previous pipeline (v4) applied two rating curves:

| Rating | Start date | Cause |
|:-------|:-----------|:------|
| Rating 1 | 2012-01-01 | Original |
| Rating 2 | 2017-07-13 | Sensor replaced; installation position may have shifted |

**Key question:** Do the new gaugings (post 2019-02-09) sit on the previous Rating 2 curve, or do they depart from it? If they sit on it, no new curve is needed and Rating 2 can simply be extended. If they depart, a new Rating 3 is required.

---

## Diagnostic -- all gaugings vs previous curves {.tabset}

### By Old/New

```{r diagnostic_old_new}
ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r1,
              aes(x = Stage_avg, y = Q_model, linetype = "Rating 1"),
              colour = "black") +
    geom_line(data = rc_v4_r2,
              aes(x = Stage_avg, y = Q_model, linetype = "Rating 2"),
              colour = "black") +
    geom_point(data = gaugings_ok,
               aes(x = Stage_avg, y = Q_meas,
                   fill = Old_New,
                   label = paste0("Event: ", Event_no,
                                  " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2, alpha = 0.8) +
    scale_linetype_manual(name = "Previous curve",
                          values = c("Rating 1" = "dashed",
                                     "Rating 2" = "solid")) +
    scale_fill_manual(values = c("Old" = "#4A90A4", "New" = "#E07B54")) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 -- all gaugings vs previous rating curves",
         subtitle = "Solid = Rating 2 (last active curve) | Dashed = Rating 1",
         fill = "Old/New",
         caption = "Old = used in previous pipeline | New = post 2019-02-09")
)
```

### By water year

```{r diagnostic_wy}
ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r1,
              aes(x = Stage_avg, y = Q_model),
              colour = "black", linetype = "dashed") +
    geom_line(data = rc_v4_r2,
              aes(x = Stage_avg, y = Q_model),
              colour = "black") +
    geom_point(data = gaugings_ok,
               aes(x = Stage_avg, y = Q_meas,
                   fill = factor(WY),
                   label = paste0("Event: ", Event_no,
                                  " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2, alpha = 0.8) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 -- all gaugings by water year vs previous rating curves",
         fill = "Water year")
)
```
```{r event_no}

ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r2,
              aes(x = Stage_avg, y = Q_model),
              colour = "black") +
    geom_point(data = gaugings_new,
               aes(x = Stage_avg, y = Q_meas,
                   colour = factor(Event_no),
                   label = paste0("Event: ", Event_no,
                                  " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 -- new gaugings by event vs Rating 2",
         colour = "Event no.")
)
```

### New gaugings only vs Rating 2

```{r diagnostic_new_only}
ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r2,
              aes(x = Stage_avg, y = Q_model),
              colour = "black") +
    geom_pointrange(data = gaugings_new,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        fill = factor(WY),
                        label = paste0("Event: ", Event_no,
                                       " | Date: ", datetime,
                                       " | Q: ", round(Q_meas, 3))),
                    shape = 21, size = 0.5) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 -- new gaugings vs previous Rating 2 curve",
         subtitle = "If points sit on the curve, no new rating is needed",
         fill = "Water year")
)
```

### Residuals -- new vs Rating 2

```{r diagnostic_resid}
cols <- c("Q_resid" = "black", "mean Q_abs_unc" = "grey", "mean Q_resid" = "red")

resid_new <- gaugings_new |>
  mutate(Stage_avg = round(Stage_avg, 1)) |>
  dplyr::select(Stage_avg, Q_meas, Q_abs_unc) |>
  inner_join(rc_v4_r2, by = "Stage_avg") |>
  mutate(resid = Q_meas - Q_model)

if (nrow(resid_new) > 0) {
  ggplot() +
    geom_line(data = resid_new,
              aes(x = Stage_avg, y = resid, colour = "Q_resid")) +
    geom_point(data = resid_new,
               aes(x = Stage_avg, y = resid, colour = "Q_resid")) +
    geom_hline(aes(yintercept =  mean(resid_new$Q_abs_unc, na.rm = TRUE),
                   colour = "mean Q_abs_unc"), linewidth = 1) +
    geom_hline(aes(yintercept = -mean(resid_new$Q_abs_unc, na.rm = TRUE),
                   colour = "mean Q_abs_unc"), linewidth = 1) +
    geom_hline(aes(yintercept =  mean(resid_new$resid, na.rm = TRUE),
                   colour = "mean Q_resid"), linewidth = 1) +
    scale_colour_manual(name = "", values = cols) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Residuals (m³/s)") +
    labs(title = "SSN844 -- residuals: new gaugings vs Rating 2",
         subtitle = paste0("n = ", nrow(resid_new), " matched gaugings | ",
                           "Mean residual = ",
                           round(mean(resid_new$resid, na.rm = TRUE), 3),
                           " m³/s | Mean uncertainty = ",
                           round(mean(resid_new$Q_abs_unc, na.rm = TRUE), 3),
                           " m³/s"))
} else {
  cat("No new gaugings matched Rating 2 stage values -- check stage rounding")
}
```

---

## Decision

_Based on the diagnostic plots above, decide whether:_

**A. Rating 2 is still valid** -- new gaugings sit within uncertainty of the old curve. No new curve needed. The previous pipeline discharge record can be extended using Rating 2.

**B. A new Rating 3 is needed** -- new gaugings systematically depart from Rating 2 beyond the measurement uncertainty. Proceed to the curve fitting section below.

---

## RC1 new curve (if needed) {.tabset}

_Complete this section only if the diagnostic shows a new curve is required._

```{r rc1_filter, include=FALSE}
# Gaugings for new curve fitting
# Using Final_rating_curve == Y and Old_New == New
# Adjust filters based on diagnostic findings

HQ_RC1_new <- gaugings_all |>
  filter(stage_status == "ok",
         Old_New == "New",
         Final_rating_curve == "Y")

message("New gaugings for RC1 fitting: ", nrow(HQ_RC1_new))
message("Stage range: ", round(min(HQ_RC1_new$Stage_avg, na.rm = TRUE), 1),
        " -- ", round(max(HQ_RC1_new$Stage_avg, na.rm = TRUE), 1), " cm")
```

_`r nrow(HQ_RC1_new)` new gaugings available for fitting (Final_rating_curve == Y, stage ok)._

### Data assessment {.tabset}

#### Overview

```{r rc1_overview}
ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r2,
              aes(x = Stage_avg, y = Q_model),
              colour = "grey50", linetype = "dashed") +
    geom_point(data = HQ_RC1_new,
               aes(x = Stage_avg, y = Q_meas,
                   fill = factor(WY),
                   label = paste0("Event: ", Event_no,
                                  " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 RC1 new -- gaugings by water year",
         subtitle = "Dashed = previous Rating 2 for reference",
         fill = "Water year")
)
```

#### By event

```{r rc1_event}
ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r2,
              aes(x = Stage_avg, y = Q_model),
              colour = "grey50", linetype = "dashed") +
    geom_point(data = HQ_RC1_new,
               aes(x = Stage_avg, y = Q_meas,
                   fill = factor(Event_no),
                   label = paste0("Event: ", Event_no,
                                  " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 RC1 new -- gaugings by event",
         fill = "Event no.")
)
```

#### Hysteresis

```{r rc1_hyst}
ggplotly(
  ggplot() +
    geom_line(data = rc_v4_r2,
              aes(x = Stage_avg, y = Q_model),
              colour = "grey50", linetype = "dashed") +
    geom_point(data = HQ_RC1_new,
               aes(x = Stage_avg, y = Q_meas,
                   fill = Stage_delta,
                   label = paste0("Event: ", Event_no,
                                  " | Date: ", datetime,
                                  " | Q: ", round(Q_meas, 3))),
               shape = 21, size = 2) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 RC1 new -- hysteresis assessment",
         fill = "Stage delta (R/F/C)")
)
```

### LOESS span selection

```{r rc1_span}
## ACTION REQUIRED: adjust span values based on data density
Rating_RC1_new <- selectspan(HQ_RC1_new, 0.15, 0.20, 0.25)

ggplotly(
  ggplot() +
    geom_pointrange(data = HQ_RC1_new,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_meas - Q_abs_unc,
                        ymax = Q_meas + Q_abs_unc,
                        label = Event_no),
                    size = 0.5) +
    geom_line(data = Rating_RC1_new,
              aes(x = Stage_avg, y = Q_model, colour = smoothed)) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 RC1 new -- LOESS span selection")
)
```

```{r rc1_span_choose, include=FALSE}
## ACTION REQUIRED: enter chosen span label and value
RC1_new_span_label <- "span_b"
RC1_new_span_value <- 0.20

Rating_RC1_new <- Rating_RC1_new |>
  filter(smoothed == RC1_new_span_label) |>
  dplyr::select(Stage_avg, Q_model)
```

### Extrapolation and final curve

```{r rc1_extrap, include=FALSE}
h0_RC1_new <- min(HQ_RC1_new$Stage_avg[HQ_RC1_new$Q_meas > 0], na.rm = TRUE) * 0.9

## ACTION REQUIRED: set inflection point for power law extrapolation
## Previous pipeline used 100cm -- check diagnostic to confirm
inflection_h_RC1_new <- 100

RC1_new_extrap_data <- HQ_RC1_new |>
  filter(Stage_avg >= inflection_h_RC1_new, Q_meas > 0)

RC1_new_loglog <- RC1_new_extrap_data |>
  mutate(log_h = log(Stage_avg - h0_RC1_new), log_Q = log(Q_meas))
lm_start_RC1_new <- lm(log_Q ~ log_h, data = RC1_new_loglog)
a_start_RC1_new  <- exp(coef(lm_start_RC1_new)[1])
b_start_RC1_new  <- coef(lm_start_RC1_new)[2]

fit_RC1_new <- tryCatch(
  nls(Q_meas ~ a * (Stage_avg - h0)^b,
      data    = RC1_new_extrap_data,
      start   = list(a = a_start_RC1_new, b = b_start_RC1_new, h0 = h0_RC1_new),
      control = nls.control(maxiter = 500)),
  error = function(e) {
    message("RC1 new nls failed with free h0 -- falling back to fixed h0")
    nls(Q_meas ~ a * (Stage_avg - h0_RC1_new)^b,
        data    = RC1_new_extrap_data,
        start   = list(a = a_start_RC1_new, b = b_start_RC1_new),
        control = nls.control(maxiter = 500))
  }
)

RC1_new_max_loess    <- max(Rating_RC1_new$Stage_avg)
RC1_new_Q_at_anchor  <- Rating_RC1_new |>
  filter(Stage_avg == RC1_new_max_loess) |>
  pull(Q_model)
RC1_new_Q_pl_anchor  <- predict(fit_RC1_new,
                                 newdata = data.frame(Stage_avg = RC1_new_max_loess))
RC1_new_pl_scale     <- RC1_new_Q_at_anchor / RC1_new_Q_pl_anchor

## ACTION REQUIRED: set max stage for extrapolation
## Previous pipeline used 146cm -- update if new higher stage observed
RC1_new_max_stage    <- 146

RC1_new_extrap_stage  <- seq(RC1_new_max_loess + 1, RC1_new_max_stage, by = 1)
RC1_new_extrap_Q      <- predict(fit_RC1_new,
                                  newdata = data.frame(Stage_avg = RC1_new_extrap_stage)) * RC1_new_pl_scale
RC1_new_extrap_values <- tibble(Stage_avg = RC1_new_extrap_stage,
                                 Q_model   = RC1_new_extrap_Q)
Rating_RC1_new_full   <- bind_rows(Rating_RC1_new, RC1_new_extrap_values) |>
  arrange(Stage_avg)

Rating_RC1_new_interp <- interp1(Rating_RC1_new_full)
names(Rating_RC1_new_interp) <- c("Stage_avg", "Q_model")

HQ_RC1_new_unc    <- HQ_unc(HQ_RC1_new, Rating_RC1_new, Rating_RC1_new_interp)
CI_input_RC1_new  <- CI_model_input(HQ_RC1_new_unc, n_boot, HQ_RC1_new)
CI_RC1_new        <- CI(CI_input_RC1_new, n_boot, RC1_new_span_value, 1)
RatingCI_RC1_new  <- bind_cols(Rating_RC1_new, CI_RC1_new)

RC1_new_extrap_CI      <- RC1_new_extrap_values |>
  mutate(Max_CI = Q_model * 1.05, Min_CI = Q_model * 0.95)
RatingCI_RC1_new_full   <- bind_rows(RatingCI_RC1_new, RC1_new_extrap_CI) |>
  arrange(Stage_avg)
RatingCI_RC1_new_interp <- interp2(RatingCI_RC1_new_full)

RC1_new_params <- coef(fit_RC1_new)
```

_h₀ = `r round(h0_RC1_new, 1)` cm | Inflection = `r round(inflection_h_RC1_new, 1)` cm | Power law anchored at LOESS endpoint (`r round(RC1_new_max_loess, 1)` cm) | a = `r round(RC1_new_params["a"], 4)`, b = `r round(RC1_new_params["b"], 4)`_

```{r rc1_extrap_plot}
ggplotly(
  ggplot() +
    geom_vline(xintercept = inflection_h_RC1_new, linetype = "dotted",
               colour = "grey50") +
    geom_line(data = Rating_RC1_new_full,
              aes(x = Stage_avg, y = Q_model,
                  colour = Stage_avg > RC1_new_max_loess)) +
    geom_point(data = HQ_RC1_new,
               aes(x = Stage_avg, y = Q_meas,
                   label = paste0("Event: ", Event_no, " | Date: ", datetime)),
               colour = ifelse(HQ_RC1_new$Stage_avg >= inflection_h_RC1_new,
                               "#E07B54", "black"),
               size = 2) +
    scale_colour_manual(values = c("FALSE" = "black", "TRUE" = "red"),
                        labels = c("LOESS", "Power law extrapolation")) +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 RC1 new -- LOESS with anchored power law extrapolation",
         subtitle = paste0("Dotted = inflection (", round(inflection_h_RC1_new, 1),
                           " cm). Orange = used for power law fit."),
         colour = "Fit type")
)
```

```{r rc1_final_plot}
Q_meas_RC1_new <- cbind(HQ_RC1_new$Stage_avg, HQ_RC1_new_unc)
colnames(Q_meas_RC1_new)[1] <- "Stage_avg"

ggplotly(
  ggplot() +
    geom_ribbon(data = RatingCI_RC1_new_interp,
                aes(x = Stage_avg, ymin = Min_CI, ymax = Max_CI),
                fill = "grey80") +
    geom_pointrange(data = Q_meas_RC1_new,
                    aes(x = Stage_avg, y = Q_meas,
                        ymin = Q_min, ymax = Q_max),
                    size = 0.5) +
    geom_line(data = RatingCI_RC1_new_interp,
              aes(x = Stage_avg, y = Q_model),
              colour = "#333333") +
    theme_bw() +
    xlab("Stage (cm)") + ylab("Discharge (m³/s)") +
    labs(title = "SSN844 -- Final rating curve v5 (new RC1)")
)
```

```{r rc1_save, include=FALSE}
## ACTION REQUIRED: update Rating number and Start date based on decision
## If Rating 2 is still valid and just extended: Rating = 2, Start = existing start
## If new curve needed: Rating = 3, Start = 2019-02-09

RC1_new_lookup <- RatingCI_RC1_new_interp |>
  mutate(
    Start  = format(as.Date("2019-02-09"), "%m/%d/%Y %H:%M"),
    Rating = 3L
  ) |>
  dplyr::select(Start, Stage_avg, Q_model, Max_CI, Min_CI, Rating)

write_csv(RC1_new_lookup,
          file.path(out_dir, "ssn844_RC1_new_rating_curve_v5.csv"))
write_csv(HQ_RC1_new,
          file.path(out_dir, "ssn844_RC1_new_stagedischarge_v5.csv"))

# Combined lookup -- append new rating to existing v4 lookup
ssn844_lookup_combined <- bind_rows(
  ratingcurve_v4 |> dplyr::select(Start, Stage_avg, Q_model, Max_CI, Min_CI, Rating),
  RC1_new_lookup
)
write_csv(ssn844_lookup_combined,
          file.path(out_dir, "ssn844_rating_curve_lookup_combined.csv"))

message("Saved: ssn844_RC1_new_rating_curve_v5.csv")
message("Saved: ssn844_rating_curve_lookup_combined.csv")
```

---

## Summary

_Complete after reviewing diagnostic and fitting decision._

| Rating | Period | n gaugings | Notes |
|:-------|:-------|:-----------|:------|
| Rating 1 (v4) | pre 2017-07-13 | — | Previous pipeline, not modified |
| Rating 2 (v4) | 2017-07-13 to 2019-02-09 | — | Previous pipeline, not modified |
| Rating 3 (new) | 2019-02-09 onward | `r nrow(HQ_RC1_new)` | New curve if diagnostic shows departure from Rating 2 |
